---
title: "Volve and H2O"
author: "Estevao Alvarenga"
date: "03/12/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
```

## Load Volve production dataset

I will use Equinor's Volve dataset, available at https://data.equinor.com/dataset/Volve (you will need to create an account for this).

The code below defines the place I have saved the production dataset.

```{r}
OD_root <- '~/OneDrive - Shell/'
file0   <- 'Volve production data.xlsx'
path0   <- paste0(OD_root,'Prodev/volve/Production_data/',file0)
```

Now I import and organise the column names in the datasets.

```{r}
prod_data_1 <- readxl::read_excel(path0,
                                  sheet = 1)
prod_data_2 <- readxl::read_excel(path0,
                                  sheet = 2,
                                  skip = 1)

colnames(prod_data_2) <- names(readxl::read_excel(path0,
                                                  sheet = 2,
                                                  n_max = 1)) %>%
  make.names()

prod_data_2 <- prod_data_2 %>%
  mutate(On.Stream = as.numeric(On.Stream),
         Oil = as.numeric(Oil),
         Gas = as.numeric(Gas),
         Water = as.numeric(Water),
         GI = as.numeric(GI),
         WI = as.numeric(WI))


head(prod_data_1)
names(prod_data_1)
str(prod_data_1)
summary(prod_data_1)

head(prod_data_2)
names(prod_data_2)
str(prod_data_2)
summary(prod_data_2)
```

Sheet 1 contains daily production data, while sheet 2 cointains monthly aggregates. Because of what I want to accomplish with this exercise, I will from now focus on the daily data.

## Exploring Volve's daily production data

Firstly, lets see the number and types of wells we have:

```{r}
prod_data_1 %>%
  select(WELL_BORE_CODE, NPD_WELL_BORE_NAME, WELL_TYPE, FLOW_KIND, BORE_WI_VOL) %>%
  distinct()
```

So I guess 15/9-F-1C and 15/9-F-5 were converted at some point, but why they keep the `FLOW_KIND` variable as `production` I still do not know. Also the `BORE_WI_VOL` is a bit odd right now.

Lets look at the volumes over time for well 15/9-F-5.

```{r}
prod_data_1 %>%
  filter(NPD_WELL_BORE_NAME == '15/9-F-5') %>%
  select(DATEPRD, NPD_WELL_BORE_NAME, WELL_TYPE,
         BORE_OIL_VOL, BORE_GAS_VOL, BORE_WAT_VOL) %>%
  gather(fluid, volume, -DATEPRD, -NPD_WELL_BORE_NAME, -WELL_TYPE) %>%
  ggplot(aes(x = DATEPRD, y = volume)) +
  geom_point() +
  facet_wrap(~fluid + WELL_TYPE, ncol = 2, scales = 'free_y')
```

I find it a bit odd to have production and water injection in the same year on the same well.

```{r}
prod_data_1_summary <- prod_data_1 %>%
  group_by(WELL_BORE_CODE, NPD_WELL_BORE_NAME, WELL_TYPE, FLOW_KIND) %>%
  summarise(number_of_rows = n(),
            min_date = min(DATEPRD),
            max_date = max(DATEPRD))
```

By looking at the count of rows on each class, I guess the WI and OP tag were misplaced. I will continue with this assumtion for now.

```{r}
prod_data_1 %>%
  group_by(NPD_WELL_BORE_NAME, FLOW_KIND) %>%
  summarise(number_of_rows = n(),
            min_date = min(DATEPRD),
            max_date = max(DATEPRD))
```

So this is a 6P2I project. Still a bit odd the two injectors start 5 months before the two first producers.

To make my plots better to visualise, I will convert all fluids to kboe.

```{r}
prod_data_1_boe <- prod_data_1 %>%
  mutate(qg_boe = BORE_GAS_VOL / 170000,
         qo_boe = BORE_OIL_VOL * 6.29,
         qw_boe = BORE_WAT_VOL * 6.29)
```

Now I will make one plot per well.

```{r}
prod_data_1_boe %>%
  select(DATEPRD, NPD_WELL_BORE_NAME,
         qg_boe, qo_boe, qw_boe) %>%
  gather(fluid, volume, -DATEPRD, -NPD_WELL_BORE_NAME) %>%
  ggplot(aes(x = DATEPRD, y = volume)) +
  geom_point(aes(colour = fluid)) +
  facet_wrap(~NPD_WELL_BORE_NAME, ncol = 1, scales = 'free_y')
```

Ok, after looking at this plot, I will go back to explore the monthly aggregates, to come up with an hypothesis of what is happening here.

First I will create similar conversions as for the daily data, in addition to calculating daily averages per month, then I will plot.
```{r}

days_in_month <- data.frame(
  Month         = 1:12,
  days_in_month = c(31,28,31,30,31,30,31,31,30,31,30,31)
)

prod_data_2_boe <- prod_data_2 %>%
  left_join(days_in_month) %>%
  mutate(qgp_boepd = Gas / 170000 / days_in_month,
         qo_boepd  = Oil * 6.29 / days_in_month,
         qwp_boepd = Water * 6.29 / days_in_month,
         qgi_boepd = GI / 170000 / days_in_month,
         qwi_boepd = WI * 6.29 / days_in_month,
         mth_yr    = as.Date(paste(Year, Month, '01', sep = '-')))

prod_data_2_boe %>%
  select(mth_yr, Wellbore.name,
         qgp_boepd, qo_boepd, qwp_boepd,
         qwi_boepd, qgi_boepd) %>%
  gather(fluid, volume, -mth_yr, -Wellbore.name) %>%
  ggplot(aes(x = mth_yr, y = volume)) +
  geom_point(aes(colour = fluid)) +
  facet_wrap(~Wellbore.name, ncol = 1, scales = 'free_y')
```

This looks a bit better :).

Now let's see how this looks aggregating on the field level.

```{r}
field_2_boe <- prod_data_2_boe %>%
  select(mth_yr,
         qgp_boepd, qo_boepd, qwp_boepd,
         qwi_boepd, qgi_boepd) %>%
  gather(fluid, volume, -mth_yr) %>%
  group_by(mth_yr, fluid) %>%
  summarise(fullfield_vol = sum(volume, na.rm = TRUE))

field_2_boe %>%
  ggplot(aes(x = mth_yr, y = fullfield_vol)) +
  geom_line(aes(colour = fluid))
```

## Defining the problem

I will try to forecast the oil production for the F-12 well based on the production and injection rates during the period the project had 2P2I.

```{r}
proj2p2i <- prod_data_2_boe %>%
  group_by(Wellbore.name) %>%
  mutate(Np = ifelse(is.na(cumsum(Oil)), 0, cumsum(Oil)),
         Gp_i = ifelse(is.na(cumsum(Gas)), 0, cumsum(Gas)) -
           ifelse(is.na(cumsum(GI)), 0, cumsum(GI)),
         Wp_i = ifelse(is.na(cumsum(Water)), 0, cumsum(Water)) -
           ifelse(is.na(cumsum(WI)), 0, cumsum(WI))) %>%
  ungroup() %>%
  filter(mth_yr < as.Date('2013-07-01')) %>%
  mutate() %>%
  select(mth_yr, Wellbore.name, On.Stream,
         qgp_boepd, qo_boepd, qwp_boepd,
         qwi_boepd, qgi_boepd,
         Np, Gp_i, Wp_i)

proj2p2i %>%
  gather(fluid, volume, -mth_yr, -Wellbore.name, -On.Stream) %>%
  ggplot(aes(x = mth_yr, y = volume)) +
  geom_point(aes(colour = fluid)) +
  facet_wrap(~Wellbore.name, ncol = 1, scales = 'free_y')
```

## Feature engineering

```{r}
proj2p2i %>%
  group_by(Wellbore.name) %>%
  mutate(Np = cumsum(qo_boepd) * 30)
  gather(fluid, volume, -mth_yr, -Wellbore.name, -On.Stream) %>%
```

